An Empirical Analysis of the Role of Energy in Economic Growth

### 2013-04-22 EconGrowth.Rnw
Now including diagonal lines on triangle graphs.

For the CES model (Equation \ref{eq:micEconCES_model}), the parameter $\sigma$ indicates whether capital stock and labor ($kl$) and energy ($e$) are substitutes ($\sigma >$ 1) or complements ($\sigma <$ 1). For countries whose value of $\sigma$ is below and to the right of a line connecting (0, 0) and (2, 1) on a graph of $CI_{95\%,\sigma}$ vs. $\sigma$, we can tell (with 95\% confidence) whether ($kl$) and ($e$) are substitutes or complements. For economies whose value of $\sigma$ is above and to the left of a line connecting (0, 0) and (2, 1), we cannot say with 95\% confidence whether ($kl$) and ($e$) are substitues or complements. Conversely, for economies whose value of $\sigma$ is below and to the right of a line connecting (0, 0) and (2, 1), we are able to say with 95\% confidence whether ($kl$) and ($e$) are substitues or complements.

The right column of graphs in Figure \ref{fig:CIvsParam_Lattice_Graph} shows $CI_{95\%,\sigma}$ vs. $\sigma$ for the three quantifications of energy (thermal energy, $q$, exergy, $x$, and useful work, $u$). We note that where we can tell one way or another, energy ($e$) appears to complement capital stock and labor ($kl$).

<<CIvsParam_Lattice_Graph, fig.width=ninePanelGraphWidth, fig.height=ninePanelGraphHeight, fig.pos="H", fig.cap="95\\% Confidence interval width vs. parameter value for key energy parameters ($\\gamma$, $\\delta$, and $\\sigma$) of the Cobb-Douglas and CES models.", eval=TRUE>>=
CIvsParamLattice(countryAbbrevScaling=0.6)
@

>>>>

For the CES model (Equation \ref{eq:micEconCES_model}), the parameter $\sigma$ indicates whether capital stock and labor ($kl$) and energy ($e$) are complements ($\sigma <$ 1) or substitutes ($\sigma >$ 1). For countries whose value of $\sigma$ is below and to the left of a line connecting (0, 2) and (1, 0) on a graph of $CI_{95\%,\sigma}$ vs. $\sigma$, we can say (with 95\% confidence) that ($kl$) and ($e$) are complements. For economies whose value of $\sigma$ is below and to the right of a line connecting (1, 0) and (2, 2), we can say (with 95\% confidence) that ($kl$) and ($e$) are substitutes. For economies whose value of $\sigma$ is above those lines, we cannot say with 95\% confidence whether ($kl$) and ($e$) are substitues or complements.

The right column of graphs in Figure \ref{fig:CIvsParam_Lattice_Graph} shows $CI_{95\%,\sigma}$ vs. $\sigma$ for the three quantifications of energy (thermal energy, $q$, exergy, $x$, and useful work, $u$). We note that energy ($e$) unambiguously complements capital stock and labor ($kl$) for most economies.

<<CIvsParam_Lattice_Graph, fig.width=ninePanelGraphWidth, fig.height=ninePanelGraphHeight, fig.pos="H", fig.cap="95\\% Confidence interval widths vs. parameter values for key energy parameters ($\\gamma$, $\\delta$, and $\\sigma$) of the Cobb-Douglas and CES models. If an economy appears in the region above a dashed line, the 95\\% confidence interval for the parameter extends to or beyond $\\gamma \\le$ 0 (cannot say with 95\\% confidence that energy is important for the Cobb-Douglas model), $\\delta \\ge$ 1 (cannot say with 95\\% confidence that energy is important for the CES model), or $\\sigma$ = 0 (cannot say with 95\\% confidence whether ($kl$) and ($e$) are complements or substitutes in the CES model).", eval=TRUE>>=
CIvsParamLattice(countryAbbrevScaling=0.6)
@
